---
title: R markdown report
author: "Alice Risely"
date: "22/01/2024"
output:
  html_document:
    toc: yes
    toc_depth: 4
    number_sections: true
    toc_float:
      collapsed: false
    theme: journal
    df_print: paged

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.height = 7, fig.width =10, time_it = TRUE, error = TRUE)
```

# Information

Title: Extensive diversification of amphibian skin micro- and mycobiomes upon rewilding, with limited inter-domain interactions or effects of dietary ꞵ-carotene 
Authors: Alice Risely, Phillip G. Byrne, David A, Hunter, Ana S. Carranco, Bethany J. Hoye, and Aimee J. Silla

# Abstract

The composition and dynamics of the skin bacterial and fungal microbiome is thought to influence host-pathogen defence. This microbial community is shaped by host captivity, diet, and microbial interactions between bacterial and fungal components. However, there remains little understanding of how specific micronutrients influence bacterial and fungal microbiome composition and their inter-domain interactions during rewilding of captive-bred animals. This study experimentally investigated the effect of dietary beta-carotene supplementation and subsequent field release on bacterial and fungal microbiome composition and dynamics using the Southern Corroboree frog (Pseudophryne corroboree) as a model system. We found large-scale diversification of bacterial communities post-release and similar diversification of fungal communities. The rewilded fungal mycobiome was more transient and demonstrated stronger temporal and micro-spatial fluctuations than the bacterial microbiome. Accounting for temporal and spatial factors, we found strong residual associations between bacterial members, yet limited evidence for inter-domain associations, suggesting that co-occurrence patterns between bacterial and fungal communities are largely a result of shared responses to the environment rather than direct interactions. Lastly, we found supplementation of dietary beta-carotene in captivity had no impact on post-release microbiome diversity, yet was associated with approximately 15% of common bacterial and fungal genera. Our research demonstrates that environmental factors play a dominant role over dietary beta-carotene supplementation in shaping microbiome diversity post-release, and suggest inter-domain interactions may also only exert a minor influence. Further research on the function and ecology of skin bacterial and fungal microbiomes will be crucial for developing strategies to support survival of endangered amphibian species.

Analyasis

## Load packages

```{r}

library(phyloseq)
library(ggplot2)
library(vegan)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(ape)
library(gridExtra)
library(ade4)
library(plyr)
library(tidyr)
library(data.table)
library(stringr)
library(ggrepel)
library(decontam)
library(ranacapa)
library(patchwork)
library(ggvenn)
library(gllvm)
library(ggvenn)
library(ggpubr)
library(viridis)
library(NetCoMi)
library(ggcorrplot)
library(igraph)
library(ggnetwork)
library(microbiomeutilities)
library(pheatmap)
library(RColorBrewer)
library(glmmTMB)
library(sjPlot)
library(wesanderson)


```

# Data import and cleaning

Note this data has gone through some rudimentary processing and filtering.

```{r}
setwd("C:/Users/risel/Dropbox/Academic projects/Frog microbiome UOW/Frogs_UOW/Longitudinal project/Analysis/Published")

frog_16S <- readRDS("corroboree_16S.RDS")
frog_ITS <- readRDS("corroboree_ITS.RDS")

frog_16S 
frog_ITS


```



## Study design

```{r}

ggplot(sample_data(frog_16S), aes(y = frog_id, x = TimePoint))+
  geom_line(aes(group = frog_id), alpha = 0.5)+
  geom_point(pch = 21, size = 2, width = 0.05, fill = "lightblue")+
  theme_bw()



```


## Filter samples with low read depth 

```{r}
frog_16S<-subset_samples(frog_16S, Seq_depth> 2000)
frog_ITS<-subset_samples(frog_ITS, Seq_depth> 1000)
```


## Change ASV names

```{r}
##### change names

## change ASV names 

taxtable<-data.frame(tax_table(frog_16S))
taxtable$Taxa<-row.names(taxtable)
head(taxtable)

taxtable$ASV<-1:nrow(taxtable)
taxtable$New_Taxa<-paste("B", "ASV", taxtable$ASV)

taxa_names(frog_16S)<-taxtable$New_Taxa

############################# fungi ###########

taxtable<-data.frame(tax_table(frog_ITS))
taxtable$Taxa<-row.names(taxtable)
head(taxtable)

taxtable$ASV<-1:nrow(taxtable)
taxtable$New_Taxa<-paste("F", "ASV", taxtable$ASV)

taxa_names(frog_ITS)<-taxtable$New_Taxa

```

## Rarefy

```{r}

### rarefaction curves ##########
### rarefaction curves ##########

rarefaction_V4<- ggrare(frog_16S, step = 500, se = TRUE)+xlim(c(0,5000))+ggtitle("Bacteria")+geom_vline(xintercept = 2000, linetype = "dashed")

rarefaction_ITS<- ggrare(frog_ITS, step = 500, se = TRUE)+xlim(c(0,5000))+ggtitle("Fungi")+geom_vline(xintercept = 1000, linetype = "dashed")


####### rarefy ####
####### rarefy ####
####### rarefy ####
####### rarefy ####

V4_rare <- rarefy_even_depth(frog_16S, sample.size = 2000)
ITS_rare <- rarefy_even_depth(frog_ITS, sample.size = 1000)


```

```{r}

rarefaction_V4+rarefaction_ITS

```

# Analysis: Effect of soft release

### VIZ: Barplots

Phylum level barplots

```{r}


###### phylum level ####
###### phylum level ####


V4_Phylum<- microbiome::aggregate_top_taxa(V4_rare, "Phylum", top = 11)

V4_melt<-psmelt(V4_Phylum)

V4_melt<-subset(V4_melt, OTU == "Proteobacteria")
V4_melt<- V4_melt %>% arrange(-Abundance)
dim(V4_melt)

sample_order<-V4_melt$Sample


V4_barplot <- speedyseq::plot_bar(V4_Phylum, fill = "Phylum")+
  geom_bar(  stat = "identity",width = 0.95)+
  scale_fill_brewer( palette = "Paired", direction = -1)+
  facet_grid(~TimePoint+location, scales = "free", space='free')+
  ggtitle("a) Bacteria")+
  theme(axis.text.x = element_blank(), axis.title.x = element_blank())+
  theme(legend.key.size = unit(0.5, "cm"))+
  labs(fill = "Bacterial Phylum")


V4_barplot$data$Sample <- factor(V4_barplot$data$Sample, levels = sample_order)


###################################################
###################################################
###################################################


ITS_Phylum<- microbiome::aggregate_top_taxa(ITS_rare, "Phylum", top = 11)

ITS_melt<-psmelt(ITS_Phylum)
ITS_melt<-subset(ITS_melt, OTU == "Basidiomycota")
ITS_melt<- ITS_melt %>% arrange(-Abundance)
dim(ITS_melt)

sample_order_ITS<-ITS_melt$Sample


ITS_barplot<-speedyseq::plot_bar(ITS_Phylum, fill = "Phylum")+
  geom_bar(  stat = "identity",width = 0.95)+
  scale_fill_brewer( palette = "Paired", direction = 1)+
  facet_grid(~TimePoint+location, scales = "free", space='free')+
  ggtitle("b) Fungi")+
  theme(axis.text.x = element_blank())+
  theme(legend.key.size = unit(0.5, "cm"))+
  labs(fill = "Fungal Phylum")

ITS_barplot$data$Sample <- factor(ITS_barplot$data$Sample, levels = sample_order_ITS)


V4_barplot /ITS_barplot
```

Genus level barplots.

```{r}

## genus level ###

# colour palette

names(wes_palettes)

pala<- wes_palette("Rushmore1", 5, type = c("discrete"))

palb<- wes_palette("GrandBudapest2", 4, type = c("discrete"))

palneon<- c("#af3dff", 	"#55ffe1",	"#ff3b94" ,	"#a6fd29" ,	"#37013a" )	

palx<-brewer.pal(12,"Paired")
pali<-brewer.pal(12,"Pastel1")

palz<-c(palx, pala, palb, palneon)

palz[14]<- "bisque1"
palz[17]<- "tomato1"


V4_genus<- microbiome::aggregate_top_taxa(V4_rare, "Genus", top = 24)


V4_barplot <- speedyseq::plot_bar(V4_genus, fill = "Genus")+
  geom_bar(  stat = "identity",width = 0.95)+
  scale_fill_manual( values = palz)+
  facet_grid(~TimePoint+location, scales = "free", space='free')+
  ggtitle("a) Bacteria")+
  theme(axis.text.x = element_blank(), axis.title.x = element_blank())+
  theme(legend.key.size = unit(0.4, "cm"))+
  labs(fill = "Bacterial genus")


# order samples

V4_barplot$data$Sample <- factor(V4_barplot$data$Sample, levels = sample_order)

###################################################
###################################################
###################################################


ITS_genus<- microbiome::aggregate_top_taxa(ITS_rare, "Genus", top = 24)


ITS_barplot<-speedyseq::plot_bar(ITS_genus, fill = "Genus")+
  geom_bar(  stat = "identity",width = 0.95)+
  scale_fill_manual( values = palz)+
  facet_grid(~TimePoint+location, scales = "free", space='free')+
  ggtitle("b) Fungi")+
  theme(axis.text.x = element_blank())+
  theme(legend.key.size = unit(0.4, "cm"))+
  labs(fill = "Fungal genus")

# order samples

ITS_barplot$data$Sample <- factor(ITS_barplot$data$Sample, levels = sample_order_ITS)

V4_barplot /ITS_barplot


```

### VIZ: Venn diagram

```{r}

## lab vs wild 

lab_ps<-subset_samples(frog_16S, TimePoint == "T1 (lab)")
lab_ps<-prune_taxa(taxa_sums(lab_ps)>0, lab_ps)

T2_ps<-subset_samples(frog_16S, TimePoint == "T2")
T2_ps<-prune_taxa(taxa_sums(T2_ps)>0, T2_ps)

T3_ps<-subset_samples(frog_16S, TimePoint == "T3")
T3_ps<-prune_taxa(taxa_sums(T3_ps)>0, T3_ps)


lab_ASVs<-taxa_names(lab_ps)
T2_ASVs<-taxa_names(T2_ps)
T3_ASVs<-taxa_names(T3_ps)


x <- list(
  lab = lab_ASVs,
  T2 = T2_ASVs,
  T3 = T3_ASVs
)


ggvenn(
  x, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF"),
  stroke_size = 0.5, set_name_size = 4)


#### ITS### 


lab_ps<-subset_samples(frog_ITS, TimePoint == "T1 (lab)")
lab_ps<-prune_taxa(taxa_sums(lab_ps)>0, lab_ps)

T2_ps<-subset_samples(frog_ITS, TimePoint == "T2")
T2_ps<-prune_taxa(taxa_sums(T2_ps)>0, T2_ps)

T3_ps<-subset_samples(frog_ITS, TimePoint == "T3")
T3_ps<-prune_taxa(taxa_sums(T3_ps)>0, T3_ps)

lab_ASVs<-taxa_names(lab_ps)
T2_ASVs<-taxa_names(T2_ps)
T3_ASVs<-taxa_names(T3_ps)


x <- list(
  lab = lab_ASVs,
  T2 = T2_ASVs,
  T3 = T3_ASVs
)


ggvenn(
  x, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF"),
  stroke_size = 0.5, set_name_size = 4)



```



### Prevalence stats - genus level

Make sure we get representative genera from across time points given that samples from lab frogs are more common. This may cause a bias in genus selection.

```{r}
# prevalence function

prevalence <- function(physeq, add_tax = TRUE){
  
  ## Check if taxa are rows
  trows <- taxa_are_rows(physeq)
  
  ## Extract OTU table
  otutab <- as.data.frame(otu_table(physeq))
  
  ## Transpose OTU table (species should be arranged by rows)
  if(trows == FALSE){
    otutab <- t(otutab)
  }
  
  ## Estimate prevalence (number of samples with OTU present)
  prevdf <- apply(X = otutab,
                  #MARGIN = ifelse(trows, yes = 1, no = 2),  # for a non-transposed data
                  MARGIN = 1,
                  FUN = function(x){sum(x > 0)})
  
  ## Add total and average read counts per OTU
  prevdf <- data.frame(Prevalence = prevdf,
                       TotalAbundance = taxa_sums(physeq),
                       MeanAbundance = rowMeans(otutab),
                       MedianAbundance = apply(otutab, 1, median))
  
  ## Add taxonomy table
  if(add_tax == TRUE && !is.null(tax_table(physeq, errorIfNULL = F))){
    prevdf <- cbind(prevdf, tax_table(physeq))
  }
  return(prevdf)
}

```


#### Select common bacterial genera to retain

```{r}

V4_genus<-tax_glom(V4_rare, "Genus")

V4_genus_T1 <-subset_samples(V4_genus, TimePoint == "T1 (lab)")
V4_genus_T2 <-subset_samples(V4_genus, TimePoint == "T2")
V4_genus_T3 <-subset_samples(V4_genus, TimePoint == "T3")

V4_prev_T1<-prevalence(V4_genus_T1)
V4_prev_T2<-prevalence(V4_genus_T2)
V4_prev_T3<-prevalence(V4_genus_T3)

V4_prev_T1<-V4_prev_T1%>%arrange(-Prevalence)
V4_prev_T2<-V4_prev_T2%>%arrange(-Prevalence)
V4_prev_T3<-V4_prev_T3%>%arrange(-Prevalence)


# top genera per group

V4_top_genera_T1<- V4_prev_T1[1:12,]
V4_top_genera_T1$Genus

V4_top_genera_T2<- V4_prev_T2[1:12,]
V4_top_genera_T2$Genus

V4_top_genera_T3<- V4_prev_T3[1:12,]
V4_top_genera_T3$Genus

# select genera to keep


V4_genera_to_keep<-unique(c(V4_top_genera_T1$Genus, V4_top_genera_T2$Genus, V4_top_genera_T3$Genus))
V4_genera_to_keep<- V4_genera_to_keep[V4_genera_to_keep != "uncultured"]
V4_genera_to_keep<- V4_genera_to_keep[V4_genera_to_keep != "uncultured bacterium"]
V4_genera_to_keep

```

#### Select common fungal genera to retain
```{r}
############## ITS #####

ITS_genus<-tax_glom(ITS_rare, "Genus")

ITS_genus_T1 <-subset_samples(ITS_genus, TimePoint == "T1 (lab)")
ITS_genus_T2 <-subset_samples(ITS_genus, TimePoint == "T2")
ITS_genus_T3 <-subset_samples(ITS_genus, TimePoint == "T3")


ITS_prev_T1<-prevalence(ITS_genus_T1)
ITS_prev_T2<-prevalence(ITS_genus_T2)
ITS_prev_T3<-prevalence(ITS_genus_T3)

ITS_prev_T1<-ITS_prev_T1%>%arrange(-Prevalence)
ITS_prev_T2<-ITS_prev_T2%>%arrange(-Prevalence)
ITS_prev_T3<-ITS_prev_T3%>%arrange(-Prevalence)


# top genera per group

ITS_top_genera_T1<- ITS_prev_T1[1:12,]
ITS_top_genera_T1$Genus

ITS_top_genera_T2<- ITS_prev_T2[1:12,]
ITS_top_genera_T2$Genus

ITS_top_genera_T3<- ITS_prev_T3[1:12,]
ITS_top_genera_T3$Genus



ITS_genera_to_keep<-unique(c(ITS_top_genera_T1$Genus, ITS_top_genera_T2$Genus, ITS_top_genera_T3$Genus))
ITS_genera_to_keep<- ITS_genera_to_keep[ITS_genera_to_keep != "uncultured"]
ITS_genera_to_keep<- ITS_genera_to_keep[ITS_genera_to_keep != "Kingdom_Fungi"]
ITS_genera_to_keep

```


### VIZ: Heat map

```{r}


# create a gradient color palette for abundance
grad_ab <- colorRampPalette(c("#faf3dd","#f7d486" ,"#5e6472"))
grad_ab <- colorRampPalette(c("mistyrose", "lightblue", "cadetblue", "navyblue"))
grad_ab_pal <- grad_ab(10)


V4_top<-subset_taxa(V4_rare, Genus %in% V4_genera_to_keep)
V4_top<-tax_glom(V4_top, "Genus")


ITS_top<-subset_taxa(ITS_rare, Genus %in% ITS_genera_to_keep)
ITS_top<-tax_glom(ITS_top, "Genus")

# Define colors for each sample type
# Specify colors

pal<-pals::stepped3()[c(1,5,9,13)]
pal[2]<-"violetred"
#pal[3]<-"aquamarine4"
pal
ann_colors = list(
  TimePoint = c(`T1 (lab)` = "#3182BD", T2 = "violetred", T3 = "#31A354"))


###############
###############
###############
###############

taxa_names(V4_top)<-paste(abbreviate(V4_top@tax_table@.Data[,2],6), 1:length(taxa_names(V4_top)), sep = "")


V4_top@tax_table@.Data[,6]<-abbreviate(V4_top@tax_table@.Data[,6],15)


heatmap1<-plot_taxa_heatmap(V4_top,
                            subset.top = 25,
                            VariableA = c("TimePoint"),
                            heatcolors = grad_ab_pal,
                            annotation_colors= ann_colors,
                            transformation = "log10",
                            cluster_rows = T,
                            cluster_cols = T,
                            clustering_method = "ward.D2",
                            show_colnames = F,
                            fontsize = 8,
                            cellwidth = 1.5, 
                            cellheight = 8)


## fungi

taxa_names(ITS_top)<-paste(abbreviate(ITS_top@tax_table@.Data[,2],6), 1:length(taxa_names(ITS_top)), sep = "")


ITS_top@tax_table@.Data[,6]<-abbreviate(ITS_top@tax_table@.Data[,6],15)


heatmap2<-plot_taxa_heatmap(ITS_top,
                            subset.top = 25,
                            VariableA = c("TimePoint"),
                            heatcolors = grad_ab_pal, 
                            annotation_colors= ann_colors,
                            transformation = "log10",
                            cluster_rows = T,
                            cluster_cols = T,
                            clustering_method = "ward.D2",
                            show_colnames = F,
                            fontsize = 8,
                            cellwidth = 1.5, 
                            cellheight = 8)


ggpubr::ggarrange(heatmap1$plot[[4]], heatmap2$plot[[4]],  ncol = 1, align = "v")



```


## STATS: GLLVM model
#### Data prep: Agglomerate, transform, filter and merge ITS and 16S data

Agglomerate to genus, CLR transform, filter and merge ITS and 16S data. Data is CLR transformed before filtering to ensure CLR values represent compositional values prior to filtering. 

```{r}


V4_samples<-sample_names(frog_16S)
ITS_samples<-sample_names(frog_ITS)
samples_to_keep<-intersect(V4_samples,ITS_samples)


frog_16S_sub<-prune_samples(samples_to_keep, frog_16S)
frog_ITS_sub<-prune_samples(samples_to_keep, frog_ITS)


# glom by genus and get list of most common taxa per dataset

V4_genus<- tax_glom(frog_16S_sub, taxrank = "Genus")
ITS_genus<- tax_glom(frog_ITS_sub, taxrank = "Genus")


#########
#########

V4_top<-microbiome::aggregate_top_taxa(V4_genus, top = 26, level = "Genus")
ITS_top<-microbiome::aggregate_top_taxa(ITS_genus, top = 26, level = "Genus")

remove<-c("Other", "uncultured", "Kingdom_Fungi")

V4_to_keep<-taxa_names(V4_top)[!taxa_names(V4_top)%in% remove]
ITS_to_keep<-taxa_names(ITS_top)[!taxa_names(ITS_top)%in% remove]

# Proportion of reads left after filtering

sum(sample_sums(subset_taxa(V4_genus, Genus %in% V4_to_keep)))/sum(sample_sums(V4_genus))
sum(sample_sums(subset_taxa(ITS_genus, Genus %in% ITS_to_keep)))/sum(sample_sums(ITS_genus))


# CLR transformation


V4_clr <-microbiome::transform(V4_genus, transform = "clr")
ITS_clr <-microbiome::transform(ITS_genus, transform = "clr")

## Keep only top 25 genera

V4_clr<-subset_taxa(V4_clr, Genus %in% V4_to_keep)
ITS_clr<-subset_taxa(ITS_clr, Genus %in% ITS_to_keep)

## change names

taxa_names(V4_clr)<- data.frame(tax_table(V4_clr))$Genus
taxa_names(ITS_clr)<- data.frame(tax_table(ITS_clr))$Genus


### make merged otu table 

V4_otu<- data.frame(otu_table(V4_clr))
ITS_otu<- data.frame(otu_table(ITS_clr))

V4_otu[1:5,1:5]
ITS_otu[1:5,1:5]

names(V4_otu)<-sample_names(V4_clr)
names(ITS_otu)<-sample_names(ITS_clr)

full_otu<-rbind(V4_otu, ITS_otu)
full_otu<-otu_table(full_otu, taxa_are_rows = T)

# make merged tax table

V4_tax<- data.frame(tax_table(V4_clr))
ITS_tax<- data.frame(tax_table(ITS_clr))

full_tax<-rbind(V4_tax, ITS_tax)
full_tax2<-tax_table(full_tax)
taxa_names(full_tax2)<- row.names(full_tax)
colnames(full_tax2) <- c("Kingdom", "Phylum", "Class", 
                         "Order", "Family", "Genus", "Species")


### make merged metadata 

metadataV4<-data.frame(sample_data(V4_clr))
metadata<-sample_data(metadataV4)

## merge

merged_ps<-phyloseq(metadata, full_otu, full_tax2)
head(tax_table(merged_ps))
tail(tax_table(merged_ps))

taxa_names(merged_ps)[1:25]<- paste("B", taxa_names(merged_ps)[1:25], sep = "_")
taxa_names(merged_ps)[26:50]<- paste("F", taxa_names(merged_ps)[26:50], sep = "_")

taxa_names(merged_ps)

merged_ps

```


#### Fit GLLVM model: Effect of soft release

```{r}


merged_ps

############# data prep ###
############# data prep ###
############# data prep ###

ys <- data.frame(t(otu_table(merged_ps)))
names(ys) <-taxa_names(merged_ps)

ys<-rescale(as.matrix(ys),  to = c(-1,1))


Xs<-data.frame(sample_data(merged_ps))


Xs$location<-factor(Xs$location)
Xs$Status<-factor(Xs$Status, levels = c("Pre-release", "Post-release"))
table(Xs$TimePoint)
str(Xs)
Xs$frog_id<-factor(Xs$frog_id)

Xs$Wild<-ifelse(Xs$TimePoint == "T1 (lab)", "Lab", "Wild")
Xs$Wild<-factor(Xs$Wild)


table(Xs$Wild)

# fit model

fit <- gllvm(ys, Xs, 
             num.lv = 2,
             formula = ~  Status, 
             row.eff = ~(1|frog_id),
             starting.val = "zero",
             family = "gaussian")

#plot(fit)
#summary(fit)
#AIC(fit)

coefplot(fit, cex.ylab = 0.7)

# null model

fit_null <- gllvm(ys,
                  num.lv = 2,
                  family = "gaussian")


############################ extract for ggploting ####
############################ extract for ggploting ####
############################ extract for ggploting ####

df<-coef(fit)

est_df<-data.frame(df$Intercept)

est_df2<-data.frame(df$Xcoef) # choose columns of interest

est_df3<-merge(est_df, est_df2, by = 0)

head(est_df3)



# order genera

row.names(est_df3)<-est_df3$Row.names
est_df3<-est_df3[colnames(ys),]

head(est_df3)

#put est_df3 into long format

names(est_df3)[1]<- "Genus"
names(est_df3)[2]<- "Intercept"

estimates_df <- gather(est_df3, Treatment, Estimate, names(est_df3)[2]:names(est_df3)[ncol(est_df3)], factor_key=TRUE)

head(estimates_df )

############# extract confindence intervals  ####
############# extract confindence intervals  ####
############# extract confindence intervals  ####
############# extract confindence intervals  ####


confint_df<-data.frame(confint(fit))
dim(confint_df)

confint_df<-rbind(confint_df[rownames(confint_df) %like% "Xcoef", ],
                  confint_df[rownames(confint_df) %like% "Intercept", ])

head(confint_df)

# add a column with correct variable level
variables<- colnames(est_df3)[3:ncol(est_df3)]
variables<-c(variables, "Intercept")
variables1<-rep(variables, nrow(est_df))
variables2<-variables1[order(match(variables1, variables))]

confint_df$Treatment<-variables2


# column with taxa names. Should be automatically in the correct order but double check

confint_df$Genus<-rep(colnames(ys), length(unique(confint_df$Treatment)))



# now have estimates, confidence intervals, aas seperate data frames, but they are in different formats. Need to massage them into one dataframe for plotting.

merged<-merge(estimates_df, confint_df, by = c("Treatment", "Genus"))


final_estimates_reduced<- merged
names(final_estimates_reduced)[4]<-"CI_lower"
names(final_estimates_reduced)[5]<-"CI_upper"

#final_estimates<- merged[,c(1,2,3,7,8)]
head(final_estimates_reduced)
unique(final_estimates_reduced$Treatment)

## add significance

final_estimates_reduced$Sig<- !data.table::between(0, final_estimates_reduced$CI_lower, final_estimates_reduced$CI_upper)


final_estimates_reduced$Genus<-factor(final_estimates_reduced$Genus)

levels(final_estimates_reduced$Treatment)

final_estimates_reduced$Treatment<-factor(final_estimates_reduced$Treatment)
levels(final_estimates_reduced$Treatment)<-c("Pre-release", "Post-release")

final_estimates_reduced2<-subset(final_estimates_reduced, Treatment == "Post-release")

final_estimates_reduced2$Genus2<-abbreviate(final_estimates_reduced2$Genus, minlength = 20 )

head(final_estimates_reduced2)

final_estimates_reduced2 <- final_estimates_reduced2  %>%
  mutate(Kingdom = ifelse(substr(Genus, 1, 1) == "B", "Bacteria", "Fungi"))

P1<-ggplot(subset(final_estimates_reduced2, Kingdom == "Bacteria"), aes(y = reorder(Genus2, Estimate), x = Estimate, col = Sig))+
  geom_point()+
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper, col = Sig), height = 0, size = 0.5)+
  geom_vline(xintercept = 0, linetype = "longdash")+
  theme_bw()+
  scale_color_manual(values = c("grey", "cadetblue"))+
  theme(axis.title.y = element_blank())+
  theme(axis.text.y = element_text(face="bold", size = 8))+
  theme(legend.position = "none")+
  theme(plot.margin=unit(c(1,0.2, 0.2, 0.6),"cm"))+
  theme( strip.background = element_blank(),
         strip.text.x = element_blank())+
  theme(axis.title.x = element_blank())

P2<-ggplot(subset(final_estimates_reduced2, Kingdom == "Fungi"), aes(y = reorder(Genus2, Estimate), x = Estimate, col = Sig))+
  geom_point()+
  # facet_wrap(~Kingdom, nrow = 2, scales = "free_y")  + 
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper, col = Sig), height = 0, size = 0.5)+
  geom_vline(xintercept = 0, linetype = "longdash")+
  theme_bw()+
  scale_color_manual(values = c("grey", "cadetblue"))+
  theme(axis.title.y = element_blank())+
  theme(axis.text.y = element_text(face="bold", size = 8))+
  theme(legend.position = "none")+
  theme(plot.margin=unit(c(1,0.2, 0.2, 0.6),"cm"))+
  theme( strip.background = element_blank(),
         strip.text.x = element_blank())+
  xlab("Estimate \n <~Pre-release vs Post-release~>")

ggarrange(P1, P2, ncol = 1, labels = c("a) Bacteria", "b) Fungi"), heights = c(1, 1.15))

```



## VIZ: Alpha diversity

#### Alpha diversity: effect of soft release

```{r}

################ alpha diversity ##############


sample_data(V4_rare)$Observed<-estimate_richness(V4_rare, split = TRUE, measures = c("Observed"))$Observed
sample_data(ITS_rare)$Observed<-estimate_richness(ITS_rare, split = TRUE, measures = c("Observed"))$Observed

sample_data(V4_rare)$Shannon<-estimate_richness(V4_rare, split = TRUE, measures = c("Shannon"))$Shannon
sample_data(ITS_rare)$Shannon<-estimate_richness(ITS_rare, split = TRUE, measures = c("Shannon"))$Shannon


pal<-pals::stepped3()[c(1,5,9,13)]
pal[2]<-"violetred"


alpha1<-ggplot(sample_data(V4_rare), aes(y = Observed, x = TimePoint, col = TimePoint, fill = TimePoint))+
  geom_boxplot(alpha = 0.5, col = "black")+
  geom_jitter( pch = 21, size = 2, width = 0.1)+
  geom_line(aes(group = frog_id), alpha = 0.6, col = "grey")+
  theme_bw()+
  xlab("Time point")+
  ylab("Bacterial ASV diversity")+
  scale_color_manual(values = pal)+
  scale_fill_manual(values = pal)


alpha2<-ggplot(sample_data(ITS_rare), aes(y = Observed, x = TimePoint, col = TimePoint, fill = TimePoint))+
  geom_boxplot( alpha = 0.5, col = "black")+
  geom_jitter( pch = 21, size = 2, width = 0.1)+
  geom_line(aes(group = frog_id), alpha = 0.6, col = "grey")+
  theme_bw()+
  xlab("Time point")+
  ylab("Fungal ASV diversity")+
  scale_color_manual(values = pal)+
  scale_fill_manual(values = pal)


ggarrange(alpha1, alpha2, ncol = 2, common.legend = T)



shannon_16S<-ggplot(sample_data(V4_rare), aes(y = Shannon, x = TimePoint))+
  geom_boxplot(fill = "skyblue")+
  geom_point()+
  geom_line(aes(group = frog_id), alpha = 0.5)+
  theme_bw()+
  ylab("Shannon diversity")

shannon_ITS<-ggplot(sample_data(ITS_rare), aes(y = Shannon, x = TimePoint))+
  geom_boxplot(fill = "skyblue")+
  geom_point()+
  geom_line(aes(group = frog_id), alpha = 0.5)+
  theme_bw()+
  ylab("Shannon diversity")

ggarrange(shannon_16S, shannon_ITS)


### correlations ####
### correlations ####
### correlations ####
### correlations ####

metadata_16S<-data.frame(sample_data(V4_rare))
metadata_16S<-subset(metadata_16S, location != "Unknown"& location != "E2" & location != "E5")

metadata_ITS<-data.frame(sample_data(ITS_rare))
metadata_ITS<-subset(metadata_ITS, location != "Unknown"& location != "E2" & location != "E5")



names(metadata_16S)
names(metadata_ITS)
alpha_merged<-merge(metadata_16S[,c(1,3,7,9,14,15)], metadata_ITS[,c(1,14,15)], by = "feature.id")

head(alpha_merged)
alpha_merged$feature.id
names(alpha_merged)

names(alpha_merged)[5]<-"Seq_depth_V4"
names(alpha_merged)[6]<-"Observed_V4"

names(alpha_merged)[7]<-"Seq_depth_ITS"
names(alpha_merged)[8]<-"Observed_ITS"

cor_plot<-ggplot(alpha_merged, aes(x = Observed_V4, y = Observed_ITS))+
  geom_line(aes(group = frog_id), col = "lightgrey")+
  geom_point( aes(fill = TimePoint), size = 3, pch = 21, col = "darkgrey")+
  scale_fill_brewer(palette = "Paired")+
  theme_bw()+
  xlab("Bacterial ASV diversity")+
  ylab("Fungal ASV diversity")+
  labs(fill = "Time point")+
  labs(color = "Time point")+
  geom_smooth(method = "lm",  aes(col = TimePoint), se = F)+
  scale_color_manual(values = pal)+
  scale_fill_manual(values = pal)


cor_plot



ggarrange(alpha1, alpha2,cor_plot, ncol = 3, common.legend = T, legend = "right", labels = c("a)", "b)", "c)"))


```

### STATS: Alpha diversity stats (GLM)

```{r}

## statistics

metadata_16S<-data.frame(sample_data(V4_rare))
names(metadata_16S)
dim(metadata_16S)


detach(package:plyr)

metadata_16S %>% group_by(Status) %>% summarize(mean = mean(Observed))

##############

center <-  function(x) {
  scaled_values <- x - mean(x)
  return(scaled_values)
}

metadata_16S$frog_id<-factor(metadata_16S$frog_id)
metadata_16S$location<-factor(metadata_16S$location)
metadata_16S$Observed_scaled<- as.numeric(center(metadata_16S$Observed))
metadata_16S$Seq_depth_scaled<- as.numeric(scale(metadata_16S$Seq_depth))


### fit models ##
### fit models ##
### fit models ##

m_alpha_16S<- glmmTMB(Observed ~ Status  + Seq_depth_scaled + (1|frog_id), data = metadata_16S)

m_alpha_16S2<- glmmTMB(Observed ~ TimePoint  +Seq_depth_scaled+ (1|frog_id), data = metadata_16S)

m_alpha_16S3<- glmmTMB(Observed ~ location  +Seq_depth_scaled+ (1|frog_id), data = metadata_16S)


sjPlot::tab_model(m_alpha_16S)
sjPlot::tab_model(m_alpha_16S2)
sjPlot::plot_model(m_alpha_16S)
sjPlot::plot_model(m_alpha_16S3)

summary(m_alpha_16S2)

## shannon ##

shannon_v4<- glmmTMB(Shannon ~ Status  + Seq_depth_scaled + (1|frog_id), data = metadata_16S)

summary(shannon_v4)


########### ITS #######
########### ITS #######
########### ITS #######
########### ITS #######

metadata_ITS<-data.frame(sample_data(ITS_rare))

metadata_ITS %>% group_by(Status) %>% summarize(mean = mean(Observed))

metadata_ITS$frog_id<-factor(metadata_ITS$frog_id)
metadata_ITS$location<-factor(metadata_ITS$location)
metadata_ITS$Observed_scaled<- as.numeric(scale(metadata_ITS$Observed))
metadata_ITS$Seq_depth_scaled<- as.numeric(scale(metadata_ITS$Seq_depth))

###### fit models ###
###### fit models ###
###### fit models ###

m_alpha_ITS<- glmmTMB(Observed ~ Status  + Seq_depth_scaled+ (1|frog_id), data = metadata_ITS)

summary(m_alpha_ITS)

sjPlot::tab_model(m_alpha_ITS)


m_alpha_ITS2<- glmmTMB(Observed ~ TimePoint  + Seq_depth_scaled+ (1|frog_id), data = metadata_ITS)

sjPlot::tab_model(m_alpha_ITS2)
plot_model(m_alpha_ITS2)

metadata_ITS$Observed_predicted<- stats::predict(m_alpha_ITS2, new_data = new_data)

## shannon

shannon_ITS<- glmmTMB(Shannon ~ Status  + Seq_depth_scaled + (1|frog_id), data = metadata_ITS)

summary(shannon_ITS)
sjPlot::tab_model(shannon_ITS)
sjPlot::tab_model(shannon_v4)

#### correlation ###
#### correlation ###
#### correlation ###
#### correlation ###


head(alpha_merged)

alpha_cor_model<- glmmTMB(Observed_V4 ~ Observed_ITS + (1|frog_id), data = alpha_merged)

sjPlot::tab_model(alpha_cor_model)

summary(alpha_cor_model)


```


## Beta diversity analysis (constrained ordination)

### STATS: Beta diversity - 16S

```{r}

#############

V4_rare_sub<- V4_rare

wunifrac_dist<-distance(V4_rare_sub, method = "wunifrac")


otutable<-data.frame(t(V4_rare_sub@otu_table@.Data))
metadata <- data.frame(sample_data(V4_rare_sub))
names(metadata)


Seq_depth <- as.numeric(scale(metadata$Seq_depth))
location <-metadata$location
time_point <-metadata$TimePoint
treatment <-metadata$treatment



final_model<-capscale(wunifrac_dist ~ 
                        time_point+
                        Seq_depth,
                      env = metadata, 
                      comm = otutable) 
final_model$CCA$rank


# anova

anova_wunifrac<-anova.cca(final_model, by="terms")
anova_wunifrac

##### enclosure ######
##### enclosure ######
##### enclosure ######



final_model2<-capscale(wunifrac_dist ~ 
                         location+
                         Seq_depth,
                       env = metadata, 
                       comm = otutable) 

# weighted unifrac

anova_wunifrac<-anova.cca(final_model2, by="terms")
anova_wunifrac

```

### VIZ: Beta diversity - 16S

```{r}

####### plot effect of time point #####

## extract data from model

final_model_df<-scores(final_model)

# extract CAP scores

vectors_df<-data.frame(final_model_df$sites)
vectors_df$feature.id<-row.names(vectors_df)


# merge with info on dominant family

sample_metadata<-data.frame(sample_data(V4_rare_sub))[,c("feature.id", "TimePoint", "frog_id")]
vectors_df<-merge(vectors_df, sample_metadata, by = "feature.id")


#### add arrows #########


centroids_df<-data.frame(final_model_df$centroids)
centroids_df
row.names(centroids_df)<-c("T1 (lab)", "T2", "T3" )


###########
###########
###########
###########

vectors_wunifrac1<-vectors_df
centroids_wunifrac1<-centroids_df

# colour palette
pal<-pals::stepped3()[c(1,5,9,13)]
pal[2]<-"violetred"


V4_timepoint<-ggplot(vectors_wunifrac1, aes(x = CAP1, y = CAP2))+
  
  stat_ellipse(geom = "polygon", aes(fill = TimePoint), level = 0.9, alpha = 0.3, size = 0.5)+
  geom_point(aes(fill =TimePoint), pch = 21, size = 3, alpha = 1, stroke = 1, col = "black")+
  
  theme_bw()+
  geom_line(aes(group = frog_id), alpha = 0.2)+
  scale_fill_manual(values = pal)+
  scale_color_manual(values = pal)+
  
  # add arrows
  
  geom_segment(data=centroids_wunifrac1, aes(x = 0, y = 0, xend = CAP1, yend = CAP2),
               arrow = arrow(length = unit(0.5, "cm"), type = "closed"), lwd = 1,  col = "black")+
  ggrepel::geom_label_repel(data=centroids_wunifrac1,  
                            alpha = 0.9, col = "black", size = 4, fill = "yellow", 
                            aes(CAP1, CAP2, label = row.names(centroids_wunifrac1)))+
  
  theme_light(base_size = 12)+
  ggtitle("a) Bacteria: Time point")+
  labs(fill = "Time point", col = "Time point")

V4_timepoint



####### plot effect of enclosure #####


## extract data from model

final_model_df<-scores(final_model2)

# extract CAP scores

vectors_df<-data.frame(final_model_df$sites)
vectors_df$feature.id<-row.names(vectors_df)


# merge with info on dominant family

sample_metadata<-data.frame(sample_data(V4_rare_sub))[,c("feature.id", "location", "frog_id")]
vectors_df<-merge(vectors_df, sample_metadata, by = "feature.id")


#### add arrows #########
#### add arrows #########
#### add arrows #########
#### add arrows #########

centroids_df<-data.frame(final_model_df$centroids)
#centroids_df<-centroids_df[1:6,]
centroids_df
centroids_df<-centroids_df[1:4,]
row.names(centroids_df)<-c("E1","E2", "E3", "Lab" )


###########
###########
###########
###########

vectors_wunifrac2<-vectors_df
centroids_wunifrac2<-centroids_df



V4_location<-ggplot(vectors_wunifrac2, aes(x = CAP1, y = CAP2))+
  
  stat_ellipse(geom = "polygon", aes(fill = location), level = 0.9, alpha = 0.3, size = 0.5)+
  geom_point(aes(fill =location), pch = 21, size = 3, alpha = 1, stroke = 1, col = "black")+
  
  theme_bw()+
  geom_line(aes(group = frog_id), alpha = 0.2)+
  
  scale_fill_viridis(discrete = TRUE, direction = -1)+
  scale_color_viridis(discrete = TRUE, direction = -1)+
  
  # add arrows
  
  geom_segment(data=centroids_wunifrac2, aes(x = 0, y = 0, xend = CAP1, yend = CAP2),
               arrow = arrow(length = unit(0.5, "cm"), type = "closed"), lwd = 1,  col = "black")+
  ggrepel::geom_label_repel(data=centroids_wunifrac2,  
                            alpha = 0.9, col = "black", size = 4, fill = "yellow", 
                            aes(CAP1, CAP2, label = row.names(centroids_wunifrac2)))+
  
  theme_light(base_size = 12)+
  ggtitle("c) Bacteria: Enclosure")+
  labs(fill = "Enclosure", col = "Enclosure")

V4_location

```

### STATS: Beta diversity - ITS


```{r}

# effect of time point 

ITS_rare_sub<- subset_samples(ITS_rare, location != "Unknown")


wunifrac_dist<-distance(ITS_rare_sub, method = "wunifrac")
otutable<-data.frame(t(ITS_rare_sub@otu_table@.Data))
metadata <- data.frame(sample_data(ITS_rare_sub))
names(metadata)


Seq_depth <- as.numeric(scale(metadata$Seq_depth))
location <-metadata$location
time_point <-metadata$TimePoint


final_model<-capscale(wunifrac_dist ~ 
                        time_point+
                        Seq_depth,
                      env = metadata, 
                      comm = otutable) 


anova_wunifrac<-anova.cca(final_model, by="terms")
anova_wunifrac

# effect of enclosure 



final_model2<-capscale(wunifrac_dist ~ 
                         location+
                         Seq_depth,
                       env = metadata, 
                       comm = otutable) 


anova_wunifrac<-anova.cca(final_model2, by="terms")
anova_wunifrac


```

### VIZ: Beta diversity - ITS


```{r}

## plot for time point ##
## plot for time point ##

## extract data from model

final_model_df<-scores(final_model)

# extract CAP scores

vectors_df<-data.frame(final_model_df$sites)
vectors_df$feature.id<-row.names(vectors_df)


# merge with info on dominant family

sample_metadata<-data.frame(sample_data(ITS_rare_sub))[,c("feature.id", "TimePoint", "frog_id")]
vectors_df<-merge(vectors_df, sample_metadata, by = "feature.id")


#### add arrows #########
#### add arrows #########


centroids_df<-data.frame(final_model_df$centroids)
#centroids_df<-centroids_df[1:6,]
centroids_df
row.names(centroids_df)<-c("T1 (lab)", "T2", "T3" )


###########

vectors_wunifrac3<-vectors_df
centroids_wunifrac3<-centroids_df



ITS_timepoint<-ggplot(vectors_wunifrac3, aes(x = CAP1, y = CAP2))+
  
  stat_ellipse(geom = "polygon", aes(fill = TimePoint), level = 0.9, alpha = 0.3, size = 0.5)+
  geom_point(aes(fill =TimePoint), pch = 21, size = 3, alpha = 1, stroke = 1, col = "black")+
  geom_line(aes(group = frog_id), alpha = 0.2)+
  
  theme_bw()+
  scale_fill_manual(values = pal)+
  scale_color_manual(values = pal)+
  
  # add arrows
  
  geom_segment(data=centroids_wunifrac3, aes(x = 0, y = 0, xend = CAP1, yend = CAP2),
               arrow = arrow(length = unit(0.5, "cm"), type = "closed"), lwd = 1,  col = "black")+
  ggrepel::geom_label_repel(data=centroids_wunifrac3,  
                            alpha = 0.9, col = "black", size = 4, fill = "yellow", 
                            aes(CAP1, CAP2, label = row.names(centroids_wunifrac3)))+
  
  theme_light(base_size = 12)+
  ggtitle("b) Fungi: Time point")+
  labs(fill = "Time point", col = "Time point")

ITS_timepoint


# plot for enclosure
# plot for enclosure
# plot for enclosure


## extract data from model

final_model_df<-scores(final_model2)

# extract CAP scores

vectors_df<-data.frame(final_model_df$sites)
vectors_df$feature.id<-row.names(vectors_df)


# merge with info on dominant family

sample_metadata<-data.frame(sample_data(ITS_rare_sub))[,c("feature.id", "location", "frog_id")]
vectors_df<-merge(vectors_df, sample_metadata, by = "feature.id")


#### add arrows #########
#### add arrows #########
#### add arrows #########
#### add arrows #########

centroids_df<-data.frame(final_model_df$centroids)
centroids_df<-centroids_df[1:4,]
centroids_df
row.names(centroids_df)<-c("E1","E2", "E3",  "Lab" )


###########
###########
###########
###########

vectors_wunifrac4<-vectors_df
centroids_wunifrac4<-centroids_df



ITS_location<-ggplot(vectors_wunifrac4, aes(x = CAP1, y = CAP2))+
  
  stat_ellipse(geom = "polygon", aes(fill = location), level = 0.9, alpha = 0.3, size = 0.5)+
  geom_point(aes(fill =location), pch = 21, size = 3, alpha = 1, stroke = 1, col = "black")+
  geom_line(aes(group = frog_id), alpha = 0.2)+
  
  theme_bw()+
  scale_fill_viridis(discrete = TRUE, direction = -1)+
  scale_color_viridis(discrete = TRUE, direction = -1)+
  
  # add arrows
  
  geom_segment(data=centroids_wunifrac4, aes(x = 0, y = 0, xend = CAP1, yend = CAP2),
               arrow = arrow(length = unit(0.5, "cm"), type = "closed"), lwd = 1,  col = "black")+
  ggrepel::geom_label_repel(data=centroids_wunifrac4,  
                            alpha = 0.9, col = "black", size = 4, fill = "yellow", 
                            aes(CAP1, CAP2, label = row.names(centroids_wunifrac4)))+
  
  theme_light(base_size = 12)+
  ggtitle("d) Fungi: Enclosure")+
  labs(fill = "Enclosure", col = "Enclosure")



# plot together
beta1<-ggarrange(V4_timepoint, ITS_timepoint, common.legend = T, legend = "right")

beta2<-ggarrange(V4_location, ITS_location, common.legend = T, legend = "right")

ggarrange(beta1, beta2, ncol = 1)

```

## STATS: Homogeneity of variance test

```{r}

# bacteria

wunifrac_dist<-distance(V4_rare_sub, method = "wunifrac")

# Extract the group membership
groups <- sample_data(V4_rare_sub)$TimePoint

# Perform beta dispersion
beta_disp <- betadisper(wunifrac_dist, groups)

# View the results
print(beta_disp)
plot(beta_disp)

# Perform an analysis of variance on the beta dispersion result
anova_beta_disp <- anova(beta_disp)
anova_beta_disp 

# fungi

wunifrac_dist<-distance(ITS_rare_sub, method = "wunifrac")

# Extract the group membership
groups <- sample_data(ITS_rare_sub)$TimePoint

# Perform beta dispersion
beta_disp <- betadisper(wunifrac_dist, groups)

# View the results
print(beta_disp)
plot(beta_disp)

# Perform an analysis of variance on the beta dispersion result
anova_beta_disp <- anova(beta_disp)
anova_beta_disp 

```

## Network analysis

#### Residual correlation plot

```{r}


merged_ps

############# data prep ###
############# data prep ###
############# data prep ###

ys <- data.frame(t(otu_table(merged_ps)))
names(ys) <-taxa_names(merged_ps)

ys<-rescale(as.matrix(ys),  to = c(-1,1))

Xs<-data.frame(sample_data(merged_ps))

Xs$location<-factor(Xs$location)
Xs$Status<-factor(Xs$Status, levels = c("Pre-release", "Post-release"))
table(Xs$TimePoint)
str(Xs)
Xs$frog_id<-factor(Xs$frog_id)

Xs$Wild<-ifelse(Xs$TimePoint == "T1 (lab)", "Lab", "Wild")
Xs$Wild<-factor(Xs$Wild)


table(Xs$Wild)

# fit model

fit <- gllvm(ys, Xs, 
             num.lv = 2,
             formula = ~  TimePoint+location, 
             starting.val = "zero",
             family = "gaussian")

coefplot(fit)

### residual correlation

cr1<-data.frame(getResidualCor(fit))#


names(cr1)<-names(data.frame(ys))
names(cr1)<-abbreviate(names(cr1), minlength = 15)

row.names(cr1)<-names(data.frame(ys))
rownames(cr1)<-abbreviate(rownames(cr1), minlength = 15 )

cr2<-cor_pmat(cr1)


# Regions coloured in dark blue in correlation plot indicate 
# clusters of species that are positively correlated with each other, 
#after controlling for (co)variation in species explained by 
#environmental terms

corplot1<-ggcorrplot(cr1, 
                     hc.order = F,
                     outline.col = "white",
                     type = "full",
                     ggtheme = ggplot2::theme_minimal,
                     tl.cex = 7.5,
                     p.mat = cr2,
                     sig.level = 0.01,
                     #show.diag = F,
                     insig = "blank",
                     #  colors = c("#6D9EC1", "white", "#E46726"))
                     colors = c("blue", "white", "red"))+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5))+
  theme(plot.margin=unit(c(1,1,1,1),"cm"))+
  geom_vline(xintercept = 25.5)+
  geom_hline(yintercept = 25.5)
# annotate("rect", xmin = 21.5, xmax = 22.5, ymin = 0, ymax = 50.5,
#  alpha = .5,fill = NA, col = "black", linetype = "dashed")

corplot1

###### null model ###
###### null model ###
###### null model ###

fit_null <- gllvm(ys,
                  num.lv = 2,
                  starting.val = "zero",
                  family = "gaussian")


# amount of variation explained by TimePoint

rcov <- getResidualCov(fit, adjust = 0)
rcov0 <- getResidualCov(fit_null, adjust = 0)
rcov0$trace; rcov$trace
1 - rcov$trace / rcov0$trace

#The ratio of traces suggests that environmental variables explain approximately 68% of the (co)variation in microbial species abundances.

cr1_null<-data.frame(getResidualCor(fit_null))#


names(cr1_null)<-names(data.frame(ys))
names(cr1_null)<-abbreviate(names(cr1_null), minlength = 15)

row.names(cr1_null)<-names(data.frame(ys))
rownames(cr1_null)<-abbreviate(rownames(cr1_null), minlength = 15 )

cr2_null<-cor_pmat(cr1_null)

#library(ggcorrplot)

# Regions coloured in dark blue in correlation plot indicate 
# clusters of species that are positively correlated with each other, 
#after controlling for (co)variation in species explained by 
#environmental terms

ggcorrplot(cr1_null, 
           hc.order = F,
           outline.col = "white",
           type = "full",
           ggtheme = ggplot2::theme_minimal,
           tl.cex = 7.5,
           p.mat = cr2_null,
           sig.level = 0.01,
           #show.diag = F,
           insig = "blank",
           #  colors = c("#6D9EC1", "white", "#E46726"))
           colors = c("blue", "white", "red"))+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5))+
  theme(plot.margin=unit(c(1,1,1,1),"cm"))+
  geom_vline(xintercept = 25.5)+
  geom_hline(yintercept = 25.5)

```

#### Network figure

```{r}
dim(cr1)

corplot2<-ggcorrplot(cr1, 
                     hc.order = F,
                     outline.col = "white",
                     type = "lower",
                     ggtheme = ggplot2::theme_minimal,
                     tl.cex = 7.5,
                     p.mat = cr2,
                     sig.level = 0.01,
                     #show.diag = F,
                     insig = "blank",
                     # colors = c("#6D9EC1", "white", "#E46726"))
                     colors = c("blue", "white", "red"))+
  # theme(axis.text.x = element_text(angle = 90, vjust=0.5))+
  theme(plot.margin=unit(c(0.5,0.5,0.5,0.5),"cm"))+
  geom_vline(xintercept = 25.5)+
  geom_hline(yintercept = 25.5)



### generate network data

edge_data<-corplot2$data


vertex_data<- data.frame(unique(c(edge_data$Var1,edge_data$Var2)))
names(vertex_data)[1]<-"name"

vertex_data$Kingdom<- ifelse(grepl('F_', vertex_data$name), "Fungi", "Bacteria")

edge_data<-subset(edge_data, signif == 1)
edge_data<-subset(edge_data, value != 1)
edge_data<-edge_data[,c(1:3)]
names(edge_data)<-c("To", "From", "cor")

edge_data$weight2 <- abs(edge_data$cor)
edge_data$dir <- ifelse(edge_data$cor<0, "Neg", "Pos")
edge_data$weight<- as.numeric(rescale(edge_data$cor, c(0.01,0.2)))


g <- graph_from_data_frame(edge_data, directed = F, vertices = vertex_data)


n<-ggnetwork::fortify(g, layout = igraph::with_fr())

head(n)

n_full<-n


network<-ggplot(n_full, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges( aes(col = dir, alpha = weight2)) +
  theme_blank()+
  geom_nodes(aes(fill = Kingdom), pch = 21,  size = 5, col = "white")+
  scale_fill_manual(values = c("gold", "deepskyblue3"))+
  scale_color_manual(values = c("blue", "red"))+
  labs(col = "Direction")+
  theme(plot.margin=unit(c(2,1,1,1),"cm"))+
  guides(alpha = "none")+
  theme(legend.key.height = unit(0.005, "cm"))+
  theme(legend.position = c(0.9,0.98))+
  theme(legend.background = element_rect(fill="lightblue",
                                         linewidth=0.5, 
                                         linetype="solid", 
                                         colour ="lightblue"))

ggarrange(corplot1, network, ncol = 2, 
          labels = c("a)", "b)"), widths = c(1.5,1))


edge_data$weight<- as.numeric(rescale(edge_data$cor, c(0.01,0.05)))


network1<-ggplot(n_full, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges( aes(col = dir, alpha = weight2)) +
  theme_blank()+
  geom_nodes(aes(fill = Kingdom), pch = 21,  size = 5, col = "white")+
  scale_fill_manual(values = c("gold", "skyblue"))+
  scale_color_manual(values = c("blue", "red"))+
  labs(col = "Direction")+
  theme(plot.margin=unit(c(2,1,1,1),"cm"))+
  guides(alpha = "none")+
  theme(legend.key.height = unit(0.005, "cm"))+
  theme(legend.position = c(0.05,0.1))+
  theme(legend.background = element_rect(fill="lightblue",
                                         linewidth=0.5, 
                                         linetype="solid", 
                                         colour ="lightblue"))+
  geom_label(aes(label = name, fill = Kingdom), size = 2)


ggarrange(corplot1, network1, ncol = 2, 
          labels = c("a)", "b)"), widths = c(1.5,1))


```



# Analysis: Effect of carotenoid treatment (Fig 7)

### Beta diversity: lab samples

```{r}
ps_lab_V4 <- subset_samples(V4_rare, TimePoint == "T1 (lab)")
ps_lab_ITS <- subset_samples(ITS_rare, TimePoint == "T1 (lab)")

```

```{r}
####

wunifrac_dist<-distance(ps_lab_V4, method = "wunifrac")
otutable<-data.frame(t(ps_lab_V4@otu_table@.Data))
metadata <- data.frame(sample_data(ps_lab_V4))
names(metadata)

ggplot(metadata, aes(y = mass, x = clutch))+geom_boxplot()+geom_point()


Seq_depth <- as.numeric(scale(metadata$Seq_depth))
treatment <-metadata$treatment
clutch <- metadata$clutch
mass<-metadata$mass
summary(mass)

names(metadata)



final_model<-capscale(wunifrac_dist ~ 
                        
                        treatment+
                        Seq_depth,
                      env = metadata, 
                      comm = otutable) 


anova_wunifrac<-anova.cca(final_model, by="terms")
anova_wunifrac

####### plot #####
####### plot #####
####### plot #####


## extract data from model

final_model_df<-scores(final_model)

# extract CAP scores

vectors_df<-data.frame(final_model_df$sites)
vectors_df$feature.id<-row.names(vectors_df)


# merge with info on dominant family

sample_metadata<-data.frame(sample_data(ps_lab_V4))[,c("feature.id", "treatment")]
vectors_df<-merge(vectors_df, sample_metadata, by = "feature.id")


#### add arrows #########
#### add arrows #########
#### add arrows #########
#### add arrows #########

centroids_df<-data.frame(final_model_df$centroids)
#centroids_df<-centroids_df[1:6,]
centroids_df


###########
###########
###########
###########

vectors_wunifrac1<-vectors_df
centroids_wunifrac1<-centroids_df

# colour palette
pal<-pals::stepped3()[c(1,5,9,13)]
#pal<-pals::tol()[c(1,3,4,12)]


p1<- ggplot(vectors_wunifrac1, aes(x = CAP1, y = CAP2))+
  
  stat_ellipse(geom = "polygon", aes(fill = treatment, col = treatment), level = 0.9, alpha = 0.3, size = 0.5)+
  geom_point(aes(fill =treatment), pch = 21, size = 3, alpha = 1, stroke = 1, col = "black")+
  theme_bw()+
  scale_fill_manual(values = pal)+
  scale_color_manual(values = pal)+
  
  theme_light(base_size = 12)+
  labs(fill = "Carot. treatment", col = "Carot. treatment")


####### ITS ##########
####### ITS ##########
####### ITS ##########
####### ITS ##########


####

wunifrac_dist<-distance(ps_lab_ITS, method = "wunifrac")
otutable<-data.frame(t(ps_lab_ITS@otu_table@.Data))
metadata <- data.frame(sample_data(ps_lab_ITS))
names(metadata)


Seq_depth <- as.numeric(scale(metadata$Seq_depth))
treatment <-metadata$treatment


final_model<-capscale(wunifrac_dist ~ 
                        treatment+
                        Seq_depth,
                      env = metadata, 
                      comm = otutable) 



anova_wunifrac<-anova.cca(final_model, by="terms")
anova_wunifrac

## extract data from model

final_model_df<-scores(final_model)

# extract CAP scores

vectors_df<-data.frame(final_model_df$sites)
vectors_df$feature.id<-row.names(vectors_df)


# merge with info on dominant family

sample_metadata<-data.frame(sample_data(ps_lab_V4))[,c("feature.id", "treatment")]
vectors_df<-merge(vectors_df, sample_metadata, by = "feature.id")


#### add arrows #########
#### add arrows #########


centroids_df<-data.frame(final_model_df$centroids)
centroids_df


###########
###########
###########
###########

vectors_wunifrac2<-vectors_df
centroids_wunifrac2<-centroids_df

pal<-pals::stepped3()[c(1,5,9,13)]


p2<- ggplot(vectors_wunifrac2, aes(x = CAP1, y = CAP2))+
  
  stat_ellipse(geom = "polygon", aes(fill = treatment, col = treatment), level = 0.9, alpha = 0.3, size = 0.5)+
  geom_point(aes(fill =treatment), pch = 21, size = 3, alpha = 1, stroke = 1, col = "black")+
  
  theme_bw()+
  scale_fill_manual(values = pal)+
  scale_color_manual(values = pal)+
  
  theme_light(base_size = 12)+
  labs(fill = "Carot. treatment", col = "Carot. treatment")



ggarrange(p1, p2, ncol = 2, common.legend = T, legend = "right", labels = c("a) Bacteria", "b) Fungi"))

```


### Alpha diversity: lab samples
```{r}


sample_data(ps_lab_V4)$Observed<-estimate_richness(ps_lab_V4, split = TRUE, measures = c("Observed"))$Observed

sample_data(ps_lab_ITS)$Observed<-estimate_richness(ps_lab_ITS, split = TRUE, measures = c("Observed"))$Observed


p3<-ggplot(sample_data(ps_lab_V4), aes(y = Observed, x = treatment))+
  geom_boxplot(fill = "skyblue")+
  geom_jitter(width = 0.2)+
  xlab("Treatment")+
  ylab("Bacterial ASV div.")+
  theme_light(base_size = 12)

p4<- ggplot(sample_data(ps_lab_ITS), aes(y = Observed, x = treatment))+
  geom_boxplot(fill = "skyblue")+
  geom_jitter(width = 0.2)+
  xlab("Treatment")+
  ylab("Fungal ASV div.")+
  theme_light(base_size = 12)


ggarrange(p3, p4, p1, p2, ncol = 2, nrow = 2, common.legend = T, labels = c("a)", "b)", "c)", "d)"), legend = "bottom")


### statistics

metadata_lab_V4<-data.frame(sample_data(ps_lab_V4))

head(metadata_lab_V4)


model_16s_lab<- glmmTMB::glmmTMB(Observed~treatment + Seq_depth, data = metadata_lab_V4)
summary(model_16s_lab)


metadata_lab_ITS<-data.frame(sample_data(ps_lab_ITS))

model_ITS_lab<- glmmTMB::glmmTMB(Observed~treatment + Seq_depth, data = metadata_lab_ITS)
summary(model_ITS_lab)



```


### Beta diversity: Rewilded samples


```{r}

ps_wild_V4 <- subset_samples(V4_rare, TimePoint != "T1 (wild)")
ps_wild_ITS <- subset_samples(ITS_rare, TimePoint != "T1 (wild)")

####

wunifrac_dist<-distance(ps_wild_V4, method = "wunifrac")
otutable<-data.frame(t(ps_wild_V4@otu_table@.Data))
metadata <- data.frame(sample_data(ps_wild_V4))
names(metadata)


Seq_depth <- as.numeric(scale(metadata$Seq_depth))
treatment <-metadata$treatment
time_point<-metadata$TimePoint



final_model<-capscale(wunifrac_dist ~ 
                        treatment+
                        Seq_depth,
                      env = metadata, 
                      comm = otutable) 


# weighted unifrac

anova_wunifrac<-anova.cca(final_model, by="terms")
anova_wunifrac

####### plot #####
####### plot #####
####### plot #####


## extract data from model

final_model_df<-scores(final_model)

# extract CAP scores

vectors_df<-data.frame(final_model_df$sites)
vectors_df$feature.id<-row.names(vectors_df)


# merge with info on dominant family

sample_metadata<-data.frame(sample_data(ps_wild_V4))[,c("feature.id", "treatment")]
vectors_df<-merge(vectors_df, sample_metadata, by = "feature.id")


#### add arrows #########
#### add arrows #########


centroids_df<-data.frame(final_model_df$centroids)
centroids_df

vectors_wunifrac1<-vectors_df
centroids_wunifrac1<-centroids_df

pal<-pals::stepped3()[c(1,5,9,13)]


p1<- ggplot(vectors_wunifrac1, aes(x = CAP1, y = CAP2))+
  
  stat_ellipse(geom = "polygon", aes(fill = treatment, col = treatment), level = 0.9, alpha = 0.3, size = 0.5)+
  geom_point(aes(fill =treatment), pch = 21, size = 3, alpha = 1, stroke = 1, col = "black")+
  
  theme_bw()+
  scale_fill_manual(values = pal)+
  scale_color_manual(values = pal)+
  
  theme_light(base_size = 12)+
  labs(fill = "Carot. treatment", col = "Carot. treatment")


####### ITS ##########
####### ITS ##########
####### ITS ##########
####### ITS ##########


####

wunifrac_dist<-distance(ps_wild_ITS, method = "wunifrac")
otutable<-data.frame(t(ps_wild_ITS@otu_table@.Data))
metadata <- data.frame(sample_data(ps_wild_ITS))
names(metadata)


Seq_depth <- as.numeric(scale(metadata$Seq_depth))
treatment <-metadata$treatment
time_point<-metadata$TimePoint




final_model<-capscale(wunifrac_dist ~ 
                        treatment+
                        Seq_depth,
                      env = metadata, 
                      comm = otutable) 



anova_wunifrac<-anova.cca(final_model, by="terms")
anova_wunifrac

## extract data from model

final_model_df<-scores(final_model)

# extract CAP scores

vectors_df<-data.frame(final_model_df$sites)
vectors_df$feature.id<-row.names(vectors_df)


# merge with info on dominant family

sample_metadata<-data.frame(sample_data(ps_wild_V4))[,c("feature.id", "treatment")]
vectors_df<-merge(vectors_df, sample_metadata, by = "feature.id")


#### add arrows #########
#### add arrows #########
#### add arrows #########
#### add arrows #########

centroids_df<-data.frame(final_model_df$centroids)
centroids_df


vectors_wunifrac2<-vectors_df
centroids_wunifrac2<-centroids_df

# colour palette
pal<-pals::stepped3()[c(1,5,9,13)]


p2<- ggplot(vectors_wunifrac2, aes(x = CAP1, y = CAP2))+
  
  stat_ellipse(geom = "polygon", aes(fill = treatment, col = treatment), level = 0.9, alpha = 0.3, size = 0.5)+
  geom_point(aes(fill =treatment), pch = 21, size = 3, alpha = 1, stroke = 1, col = "black")+
  
  theme_bw()+
  scale_fill_manual(values = pal)+
  scale_color_manual(values = pal)+
  
  theme_light(base_size = 12)+
  labs(fill = "Carot. treatment", col = "Carot. treatment")



ggarrange(p1, p2, ncol = 2, common.legend = T, legend = "right", labels = c("a) Bacteria", "b) Fungi"))

```


### Alpha diversity: Rewilded samples

```{r}

##### alpha diversity ###########
##### alpha diversity ###########
##### alpha diversity ###########
##### alpha diversity ###########

sample_data(ps_wild_V4)$Observed<-estimate_richness(ps_wild_V4, split = TRUE, measures = c("Observed"))$Observed

sample_data(ps_wild_ITS)$Observed<-estimate_richness(ps_wild_ITS, split = TRUE, measures = c("Observed"))$Observed


p3<-ggplot(sample_data(ps_wild_V4), aes(y = Observed, x = treatment))+
  geom_boxplot(fill = "skyblue")+
  geom_jitter(width = 0.2)+
  xlab("Treatment")+
  ylab("Bacterial ASV div.")+
  theme_light(base_size = 12)

p4<- ggplot(sample_data(ps_wild_ITS), aes(y = Observed, x = treatment))+
  geom_boxplot(fill = "skyblue")+
  geom_jitter(width = 0.2)+
  xlab("Treatment")+
  ylab("Fungal ASV div.")+
  theme_light(base_size = 12)

ggarrange(p3, p4, p1, p2, ncol = 2, nrow = 2, common.legend = T, labels = c("e)", "f)", "g)", "h)"), legend = "bottom")


### statistics

metadata_wild_V4<-data.frame(sample_data(ps_wild_V4))

head(metadata_wild_V4)

hist(metadata_wild_V4$Observed)

model_16s_wild<- glmmTMB::glmmTMB(Observed~treatment + Seq_depth, data = metadata_wild_V4)
summary(model_16s_wild)
data.frame(broom.mixed::tidy(model_16s_wild))


metadata_wild_ITS<-data.frame(sample_data(ps_wild_ITS))

model_ITS_wild<- glmmTMB::glmmTMB(Observed~treatment + Seq_depth, data = metadata_wild_ITS)
summary(model_ITS_wild)


## Table S1

m1<- data.frame(broom.mixed::tidy(model_16s_lab))
m2<- data.frame(broom.mixed::tidy(model_ITS_lab))
m3<- data.frame(broom.mixed::tidy(model_16s_wild))
m4<- data.frame(broom.mixed::tidy(model_ITS_wild))

tableS1<-rbind(m1, m2, m3, m4)
write.table(tableS1, "TableS1.csv")


```

## GLLVM: The effect of beta-carotene of common genera

We tested the effect of beta carotene on lab and wild samples separately. To do this, we first need CKR transform data and then merge bacterial and fungal data for 1) lab samples and 2) rewilded samples.


### Data prep: Merge bacterial and fungal data for laboratory samples

```{r}

V4_samples<-sample_names(frog_16S)
ITS_samples<-sample_names(frog_ITS)
samples_to_keep<-intersect(V4_samples,ITS_samples)

#frog_ITS<- frog_ITS %>% speedyseq::orient_taxa(as = "rows")


frog_16S_sub<-prune_samples(samples_to_keep, frog_16S)
frog_ITS_sub<-prune_samples(samples_to_keep, frog_ITS)

# glom by genus and get list of most common taxa per dataset

V4_genus<- tax_glom(frog_16S_sub, taxrank = "Genus")
ITS_genus<- tax_glom(frog_ITS_sub, taxrank = "Genus")

### filter

V4_genus<- subset_samples(V4_genus, Status == "Pre-release")
ITS_genus<- subset_samples(ITS_genus, Status == "Pre-release")


#########
#########


V4_top<-microbiome::aggregate_top_taxa(V4_genus, top = 42, level = "Genus")
ITS_top<-microbiome::aggregate_top_taxa(ITS_genus, top = 41, level = "Genus")

remove<-c("Other", "uncultured", "Kingdom_Fungi")

V4_to_keep<-taxa_names(V4_top)[!taxa_names(V4_top)%in% remove]
ITS_to_keep<-taxa_names(ITS_top)[!taxa_names(ITS_top)%in% remove]



#### CLR transform ###
#### CLR transform ###
#### CLR transform ###
#### CLR transform ###


V4_clr <-microbiome::transform(V4_genus, transform = "clr")
ITS_clr <-microbiome::transform(ITS_genus, transform = "clr")

## filter taxa

V4_clr<-subset_taxa(V4_clr, Genus %in% V4_to_keep)
ITS_clr<-subset_taxa(ITS_clr, Genus %in% ITS_to_keep)

## change names

taxa_names(V4_clr)<- data.frame(tax_table(V4_clr))$Genus
taxa_names(ITS_clr)<- data.frame(tax_table(ITS_clr))$Genus


### make merged otu table 

V4_otu<- data.frame(otu_table(V4_clr))
ITS_otu<- data.frame(otu_table(ITS_clr))

V4_otu[1:5,1:5]
ITS_otu[1:5,1:5]

names(V4_otu)<-sample_names(V4_clr)
names(ITS_otu)<-sample_names(ITS_clr)

full_otu<-rbind(V4_otu, ITS_otu)
full_otu<-otu_table(full_otu, taxa_are_rows = T)

# make merged tax table

V4_tax<- data.frame(tax_table(V4_clr))
ITS_tax<- data.frame(tax_table(ITS_clr))

full_tax<-rbind(V4_tax, ITS_tax)
full_tax2<-tax_table(full_tax)
taxa_names(full_tax2)<- row.names(full_tax)
colnames(full_tax2) <- c("Kingdom", "Phylum", "Class", 
                         "Order", "Family", "Genus", "Species")


### make merged metadata 

metadataV4<-data.frame(sample_data(V4_clr))
metadata<-sample_data(metadataV4)

## merge

merged_ps<-phyloseq(metadata, full_otu, full_tax2)
head(tax_table(merged_ps))
tail(tax_table(merged_ps))

taxa_names(merged_ps)[1:40]<- paste("B", taxa_names(merged_ps)[1:40], sep = "_")
taxa_names(merged_ps)[41:80]<- paste("F", taxa_names(merged_ps)[41:80], sep = "_")

merged_lab_ps<-merged_ps

merged_lab_ps


```

### Data prep: Merge bacterial and fungal data for rewilded samples

```{r}

V4_samples<-sample_names(frog_16S)
ITS_samples<-sample_names(frog_ITS)
samples_to_keep<-intersect(V4_samples,ITS_samples)


frog_16S_sub<-prune_samples(samples_to_keep, frog_16S)
frog_ITS_sub<-prune_samples(samples_to_keep, frog_ITS)

# glom by genus and get list of most common taxa per dataset

V4_genus<- tax_glom(frog_16S_sub, taxrank = "Genus")
ITS_genus<- tax_glom(frog_ITS_sub, taxrank = "Genus")

### filter

V4_genus<- subset_samples(V4_genus, Status == "Post-release")
ITS_genus<- subset_samples(ITS_genus, Status == "Post-release")


#########
#########
#########
#########

V4_top<-microbiome::aggregate_top_taxa(V4_genus, top = 46, level = "Genus")
ITS_top<-microbiome::aggregate_top_taxa(ITS_genus, top = 41, level = "Genus")

remove<-c("Other", "uncultured", "Kingdom_Fungi", "uncultured bacterium")

V4_to_keep<-taxa_names(V4_top)[!taxa_names(V4_top)%in% remove]
ITS_to_keep<-taxa_names(ITS_top)[!taxa_names(ITS_top)%in% remove]

#### CLR transform ###
#### CLR transform ###
#### CLR transform ###
#### CLR transform ###


V4_clr <-microbiome::transform(V4_genus, transform = "clr")
ITS_clr <-microbiome::transform(ITS_genus, transform = "clr")

## filter taxa

V4_clr<-subset_taxa(V4_clr, Genus %in% V4_to_keep)
ITS_clr<-subset_taxa(ITS_clr, Genus %in% ITS_to_keep)

## change names

taxa_names(V4_clr)<- data.frame(tax_table(V4_clr))$Genus
taxa_names(ITS_clr)<- data.frame(tax_table(ITS_clr))$Genus


### make merged otu table 

V4_otu<- data.frame(otu_table(V4_clr))
ITS_otu<- data.frame(otu_table(ITS_clr))

V4_otu[1:5,1:5]
ITS_otu[1:5,1:5]

names(V4_otu)<-sample_names(V4_clr)
names(ITS_otu)<-sample_names(ITS_clr)

full_otu<-rbind(V4_otu, ITS_otu)
full_otu<-otu_table(full_otu, taxa_are_rows = T)

# make merged tax table

V4_tax<- data.frame(tax_table(V4_clr))
ITS_tax<- data.frame(tax_table(ITS_clr))

full_tax<-rbind(V4_tax, ITS_tax)
full_tax2<-tax_table(full_tax)
taxa_names(full_tax2)<- row.names(full_tax)
colnames(full_tax2) <- c("Kingdom", "Phylum", "Class", 
                         "Order", "Family", "Genus", "Species")


### make merged metadata 

metadataV4<-data.frame(sample_data(V4_clr))
metadata<-sample_data(metadataV4)

## merge

merged_ps<-phyloseq(metadata, full_otu, full_tax2)
head(tax_table(merged_ps))
tail(tax_table(merged_ps))

taxa_names(merged_ps)[1:40]<- paste("B", taxa_names(merged_ps)[1:40], sep = "_")
taxa_names(merged_ps)[41:80]<- paste("F", taxa_names(merged_ps)[41:80], sep = "_")

merged_wild_ps<-merged_ps

merged_wild_ps

# shared taxa

table(taxa_names(merged_wild_ps) %in% taxa_names(merged_lab_ps))

taxa_names(merged_wild_ps)[taxa_names(merged_wild_ps) %in% taxa_names(merged_lab_ps)]



```


### Fit GLLVM: Lab samples

```{r}

ys <- data.frame(t(otu_table(merged_lab_ps)))
names(ys) <-taxa_names(merged_lab_ps)

ys<-rescale(as.matrix(ys),  to = c(-1,1))

Xs<-data.frame(sample_data(merged_lab_ps))


Xs$location<-factor(Xs$location)
table(Xs$TimePoint)
str(Xs)
Xs$frog_id<-factor(Xs$frog_id)
Xs$clutch<-factor(Xs$clutch)
Xs$mass_cat<-factor(Xs$mass_cat)
# fit model


fit_lab <- gllvm(ys, Xs, 
                 num.lv = 2,
                 formula = ~  treatment,
                 # row.eff = ~(1|clutch),
                 starting.val = "zero",
                 family = "gaussian")

coefplot(fit_lab, cex.ylab = 0.7)

```

#### Extract estimates and plot

```{r}

df<-coef(fit_lab)

est_df<-data.frame(df$Intercept)

est_df2<-data.frame(df$Xcoef) # choose columns of interest

est_df3<-merge(est_df, est_df2, by = 0)

head(est_df3)



# order genera

row.names(est_df3)<-est_df3$Row.names
est_df3<-est_df3[colnames(ys),]

#put est_df3 into long format

names(est_df3)[1]<- "Genus"
names(est_df3)[2]<- "Intercept"

estimates_df <- gather(est_df3, Treatment, Estimate, names(est_df3)[2]:names(est_df3)[ncol(est_df3)], factor_key=TRUE)

############# extract confindence intervals  ####
############# extract confindence intervals  ####
############# extract confindence intervals  ####
############# extract confindence intervals  ####


confint_df<-data.frame(confint(fit_lab))


confint_df<-rbind(confint_df[rownames(confint_df) %like% "Xcoef", ],
                  confint_df[rownames(confint_df) %like% "Intercept", ])

head(confint_df)

# add a column with correct variable level
variables<- colnames(est_df3)[3:ncol(est_df3)]
variables<-c(variables, "Intercept")
variables1<-rep(variables, nrow(est_df))
variables2<-variables1[order(match(variables1, variables))]

#confint_df$Treatment<-c(rep("UU", 40), rep("CU", 40), rep("UC", 40), rep("CC", 40))
confint_df$Treatment<-variables2


# column with taxa names. Should be automatically in the correct order but double check

confint_df$Genus<-rep(colnames(ys), length(unique(confint_df$Treatment)))



# now have estimates, confidence intervals, aas seperate data frames, but they are in different formats. Need to massage them into one dataframe for plotting.

merged<-merge(estimates_df, confint_df, by = c("Treatment", "Genus"))


final_estimates_reduced<- merged
names(final_estimates_reduced)[4]<-"CI_lower"
names(final_estimates_reduced)[5]<-"CI_upper"

#final_estimates<- merged[,c(1,2,3,7,8)]
head(final_estimates_reduced)
unique(final_estimates_reduced$Treatment)

final_estimates_reduced2<-final_estimates_reduced

## add significance

final_estimates_reduced2$Sig<- !data.table::between(0, final_estimates_reduced2$CI_lower, final_estimates_reduced2$CI_upper)


final_estimates_reduced2$Genus<-factor(final_estimates_reduced2$Genus)

levels(final_estimates_reduced2$Treatment)

final_estimates_reduced2$Treatment<-factor(final_estimates_reduced2$Treatment)
levels(final_estimates_reduced2$Treatment)<-c("C0", "C1", "C2", "C3")



ggplot(subset(final_estimates_reduced2, Treatment != "C0"), aes(y = reorder(Genus, Estimate), x = Estimate, col = Sig))+
  geom_point()+
  facet_wrap(~Treatment, nrow = 1, scales = "free_x")  + 
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper, col = Sig), height = 0, size = 0.5)+
  geom_vline(xintercept = 0, linetype = "longdash")+
  theme_bw()+
  scale_color_manual(values = c("grey", "red"))+
  theme(axis.title.y = element_blank())+
  theme(axis.text.y = element_text(face="bold", size = 8))+
  theme(legend.position = "none")+
  theme(plot.margin=unit(c(0.2,0.2, 0.2, 0.6),"cm"))

ggplot(subset(final_estimates_reduced2, Treatment != "C0"), aes(y = Estimate, x = Treatment, col = Sig))+
  geom_point()+
  facet_wrap(~Genus, scales = "free_y")  + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper, col = Sig), width = 0, size = 0.5)+
  geom_hline(yintercept = 0, linetype = "longdash")+
  theme_bw()+
  scale_color_manual(values = c("grey", "red"))+
  theme(axis.title.y = element_blank())+
  theme(axis.text.y = element_text(face="bold", size = 8))+
  theme(legend.position = "none")+
  theme(plot.margin=unit(c(0.2,0.2, 0.2, 0.6),"cm"))



## plot only significant

to_keep<- as.character(unique(subset(final_estimates_reduced2, Treatment != "C0" & Sig == T)$Genus))

estimates_sig<- subset(final_estimates_reduced2,  Genus %in% to_keep)



estimates_sig<-estimates_sig%>%arrange(Treatment, Genus)


estimates_sig$Genus<-factor(estimates_sig$Genus)


estimates_sig_lab<-estimates_sig

P5<-ggplot(subset(estimates_sig_lab, Treatment != "C0"), aes(y = reorder(Genus, Estimate), x = Estimate, col = Sig))+
  geom_point()+
  facet_wrap(~Treatment, nrow = 1)  + 
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper, col = Sig), height = 0, size = 0.5)+
  geom_vline(xintercept = 0, linetype = "longdash")+
  theme_bw()+
  scale_color_manual(values = c("grey", "red"))+
  theme(axis.title.y = element_blank())+
  theme(axis.text.y = element_text(face="bold", size = 12))+
  theme(legend.position = "none")+
  theme(plot.margin=unit(c(0.2,0.2, 0.2, 0.6),"cm"))

ggplot(subset(estimates_sig_lab, Genus != "B_Arthrobacter"), aes(y = Estimate, x = Treatment, col = Sig))+
  geom_point()+
  facet_wrap(~Genus)  + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper, col = Sig), width = 0, size = 0.5)+
  geom_hline(yintercept = 0, linetype = "longdash")+
  theme_bw()+
  scale_color_manual(values = c("grey", "red"))+
  theme(axis.title.y = element_blank())+
  theme(axis.text.y = element_text(face="bold", size = 10))+
  theme(legend.position = "none")+
  theme(plot.margin=unit(c(0.2,0.2, 0.2, 0.6),"cm"))




```

### Fit GLLVM: Rewilded samples

```{r}


ys <- data.frame(t(otu_table(merged_wild_ps)))
names(ys) <-taxa_names(merged_wild_ps)

ys<-rescale(as.matrix(ys),  to = c(-1,1))


Xs<-data.frame(sample_data(merged_wild_ps))


Xs$location<-factor(Xs$location)
table(Xs$TimePoint)
str(Xs)
Xs$frog_id<-factor(Xs$frog_id)

# fit model


fit_wild <- gllvm(ys, Xs, 
                  num.lv = 2,
                  formula = ~  treatment+TimePoint, 
                  starting.val = "zero",
                  family = "gaussian")

coefplot(fit_wild, cex.ylab = 0.7)




```

#### Extract estimates and plot

```{r}

df<-coef(fit_wild)

est_df<-data.frame(df$Intercept)

est_df2<-data.frame(df$Xcoef) # choose columns of interest

est_df3<-merge(est_df, est_df2, by = 0)

head(est_df3)



# order genera

row.names(est_df3)<-est_df3$Row.names
est_df3<-est_df3[colnames(ys),]

#put est_df3 into long format

names(est_df3)[1]<- "Genus"
names(est_df3)[2]<- "Intercept"

estimates_df <- gather(est_df3, Treatment, Estimate, names(est_df3)[2]:names(est_df3)[ncol(est_df3)], factor_key=TRUE)

############# extract confindence intervals  ####
############# extract confindence intervals  ####
############# extract confindence intervals  ####
############# extract confindence intervals  ####


confint_df<-data.frame(confint(fit_wild))


confint_df<-rbind(confint_df[rownames(confint_df) %like% "Xcoef", ],
                  confint_df[rownames(confint_df) %like% "Intercept", ])

head(confint_df)

# add a column with correct variable level
variables<- colnames(est_df3)[3:ncol(est_df3)]
variables<-c(variables, "Intercept")
variables1<-rep(variables, nrow(est_df))
variables2<-variables1[order(match(variables1, variables))]

#confint_df$Treatment<-c(rep("UU", 40), rep("CU", 40), rep("UC", 40), rep("CC", 40))
confint_df$Treatment<-variables2


# column with taxa names. Should be automatically in the correct order but double check

confint_df$Genus<-rep(colnames(ys), length(unique(confint_df$Treatment)))



# now have estimates, confidence intervals, aas seperate data frames, but they are in different formats. Need to massage them into one dataframe for plotting.

merged<-merge(estimates_df, confint_df, by = c("Treatment", "Genus"))


final_estimates_reduced<- merged
names(final_estimates_reduced)[4]<-"CI_lower"
names(final_estimates_reduced)[5]<-"CI_upper"

#final_estimates<- merged[,c(1,2,3,7,8)]
head(final_estimates_reduced)
unique(final_estimates_reduced$Treatment)

final_estimates_reduced2<-final_estimates_reduced

## add significance

final_estimates_reduced2$Sig<- !data.table::between(0, final_estimates_reduced2$CI_lower, final_estimates_reduced2$CI_upper)


final_estimates_reduced2$Genus<-factor(final_estimates_reduced2$Genus)

levels(final_estimates_reduced2$Treatment)

final_estimates_reduced2$Treatment<-factor(final_estimates_reduced2$Treatment)
levels(final_estimates_reduced2$Treatment)<-c("C0", "C1", "C2", "C3", "T3")

final_estimates_reduced2<-subset(final_estimates_reduced2, Treatment != "T3" )

ggplot(subset(final_estimates_reduced2, Treatment != "C0" ), aes(y = reorder(Genus, Estimate), x = Estimate, col = Sig))+
  geom_point()+
  facet_wrap(~Treatment, nrow = 1, scales = "free_x")  + 
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper, col = Sig), height = 0, size = 0.5)+
  geom_vline(xintercept = 0, linetype = "longdash")+
  theme_bw()+
  scale_color_manual(values = c("grey", "red"))+
  theme(axis.title.y = element_blank())+
  theme(axis.text.y = element_text(face="bold", size = 8))+
  theme(legend.position = "none")+
  theme(plot.margin=unit(c(0.2,0.2, 0.2, 0.6),"cm"))

ggplot(subset(final_estimates_reduced2, Treatment != "C0" & Treatment != "T3"), aes(y = Estimate, x = Treatment, col = Sig))+
  geom_point()+
  facet_wrap(~Genus, scales = "free_y")  + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper, col = Sig), width = 0, size = 0.5)+
  geom_hline(yintercept = 0, linetype = "longdash")+
  theme_bw()+
  scale_color_manual(values = c("grey", "red"))+
  theme(axis.title.y = element_blank())+
  theme(axis.text.y = element_text(face="bold", size = 8))+
  theme(legend.position = "none")+
  theme(plot.margin=unit(c(0.2,0.2, 0.2, 0.6),"cm"))



## plot only significant

to_keep<- as.character(unique(subset(final_estimates_reduced2, Treatment != "C0" & Sig == T )$Genus))

estimates_sig<- subset(final_estimates_reduced2,  Genus %in% to_keep)



estimates_sig<-estimates_sig%>%arrange(Treatment, Genus)


estimates_sig$Genus<-factor(estimates_sig$Genus)


estimates_sig_wild<-estimates_sig

P6<-ggplot(subset(estimates_sig_wild, Treatment != "C0"), aes(y = reorder(Genus, Estimate), x = Estimate, col = Sig))+
  geom_point()+
  facet_wrap(~Treatment, nrow = 1)  + 
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper, col = Sig), height = 0, size = 0.5)+
  geom_vline(xintercept = 0, linetype = "longdash")+
  theme_bw()+
  scale_color_manual(values = c("grey", "red"))+
  theme(axis.title.y = element_blank())+
  theme(axis.text.y = element_text(face="bold", size = 12))+
  theme(legend.position = "none")+
  theme(plot.margin=unit(c(0.2,0.2, 0.2, 0.6),"cm"))

ggplot(subset(estimates_sig_wild, Treatment != "UU"), aes(y = Estimate, x = Treatment, col = Sig))+
  geom_point()+
  facet_wrap(~Genus)  + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper, col = Sig), width = 0, size = 0.5)+
  geom_hline(yintercept = 0, linetype = "longdash")+
  theme_bw()+
  scale_color_manual(values = c("grey", "red"))+
  theme(axis.title.y = element_blank())+
  theme(axis.text.y = element_text(face="bold", size = 8))+
  theme(legend.position = "none")+
  theme(plot.margin=unit(c(0.2,0.2, 0.2, 0.6),"cm"))

```

## Figure 7

```{r}

ggarrange(p3, p4, p1, p2, P5, P6, ncol = 2, nrow = 3, common.legend = T, legend = "bottom")




```

